\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2018-05-04 final_JSS_edits/paper.tex
almost there with final jss edits

\Address{
  Marcus W. Beck\\
  US Environmental Protection Agency\\
  National Health and Environmental Effects Research Laboratory\\
  Gulf Ecology Division, 1 Sabine Island Drive\\
  Gulf Breeze, Florida, 32561, USA\\
  Telephone: +1/850/934-2480\\
  E-mail: \email{beck.marcus@epa.gov}
}

@@

The \pkg{RSNNS} package provides algorithms to prune connections or
nodes in a neural network \citep{Bergmeir12}.  This approach can
remove connection weights between layers or input nodes that do not
contribute to the predictive performance of the network.  In addition
to visualizing connections in the network that are not important,
connections that are pruned can be removed in successive model
fitting.  This reduces the number of free parameters (weights) that
are estimated by the model optimization algorithm, increasing the
likelihood of convergence to an estimable numeric solution for the
remaining connection weights that minimizes prediction error
\citep[i.e., model identifiability;][]{Ellenius00}. Algorithms in
\pkg{RSNNS} for weight pruning include magnitude-based pruning,
optimal brain damage, and optimal brain surgeon, whereas algorithms
for node pruning include skeletonization and the non-contributing
units method \citep{Zell98}.  The \code{plotnet} function can plot a
pruned neural network, with options to omit or display the pruned
connections (\cref{fig:plotprune}).
%
\begin{Schunk}
\begin{Sinput}
R> pruneFuncParams <- list(max_pr_error_increase = 10.0,
+    pr_accepted_error = 1.0, no_of_pr_retrain_cycles = 1000, 
+    min_error_to_stop = 0.01, init_matrix_value = 1e-6,  
+    input_pruning = TRUE, hidden_pruning = TRUE)
R> mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
+    pruneFuncParams = pruneFuncParams)
R> plotnet(mod, rel_rsc = c(3, 8))
R> plotnet(mod, prune_col = "lightblue", rel_rsc = c(3, 8))
\end{Sinput}
\end{Schunk}
%
\begin{figure}[t!]
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=1, trim = 0 20 0 65, clip]{plotprune.pdf}
\label{fig:plotprune1}
}
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=2, trim = 0 20 0 65, clip]{plotprune.pdf}
\label{fig:plotprune2}
}
\caption{A pruned neural network from \pkg{RSNNS} \citep{Bergmeir12}
  using the ``optimal brain surgeon'' algorithm described in
  \citet{Zell98}.  The default plotting behavior of \code{plotnet} is
  to omit pruned connections (\ref{fig:plotprune1}), whereas they can
  be viewed as dashed lines by including the \code{prune\_col}
  argument (\ref{fig:plotprune2}).}
\label{fig:plotprune}
\end{figure}

>>>>

\Address{
  Marcus W. Beck\\
  US Environmental Protection Agency\\
  National Health and Environmental Effects Research Laboratory\\
  Gulf Ecology Division, 1 Sabine Island Drive\\
  Gulf Breeze, Florida, 32561, USA\\
  \textit{Current address:} Southern California Coastal Water Research Project\\
  3535 Harbor Blvd., Suite 110\\
  Costa Mesa, CA, 92626\\
  Telephone: +1/714/755/3217\\
  E-mail: \email{marcusb@sccwrp.org}
}

@@

The \pkg{RSNNS} package provides algorithms to prune connections or
nodes in a neural network \citep{Bergmeir12}.  This approach can
remove connection weights between layers or input nodes that do not
contribute to the predictive performance of the network.  In addition
to visualizing connections in the network that are not important,
connections that are pruned can be removed in successive model
fitting.  This reduces the number of free parameters (weights) that
are estimated by the model optimization algorithm, increasing the
likelihood of convergence to an estimable numeric solution for the
remaining connection weights that minimizes prediction error
\citep[i.e., model identifiability;][]{Ellenius00}. Algorithms in
\pkg{RSNNS} for weight pruning include magnitude-based pruning,
optimal brain damage, and optimal brain surgeon, whereas algorithms
for node pruning include skeletonization and the non-contributing
units method \citep{Zell98}.  The \code{plotnet} function can plot a
pruned neural network, with options to omit or display the pruned
connections (\cref{fig:plotprune}).
%
\begin{Schunk}
\begin{Sinput}
R> pruneFuncParams <- list(max_pr_error_increase = 10.0,
+    pr_accepted_error = 1.0, no_of_pr_retrain_cycles = 1000, 
+    min_error_to_stop = 0.01, init_matrix_value = 1e-6,  
+    input_pruning = TRUE, hidden_pruning = TRUE)
R> mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
+    pruneFuncParams = pruneFuncParams)
R> plotnet(mod, rel_rsc = c(3, 8))
R> plotnet(mod, prune_col = "lightblue", rel_rsc = c(3, 8))
\end{Sinput}
\end{Schunk}
%
Note that the pruned network obtained with \pkg{RSNNS} and thus this
plot might vary depending on the platform used.
%
\begin{figure}[t!]
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=1, trim = 0 20 0 65, clip]{plotprune.pdf}
\label{fig:plotprune1}
}
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=2, trim = 0 20 0 65, clip]{plotprune.pdf}
\label{fig:plotprune2}
}
\caption{A pruned neural network from \pkg{RSNNS} \citep{Bergmeir12}
  using the ``optimal brain surgeon'' algorithm described in
  \citet{Zell98}.  The default plotting behavior of \code{plotnet} is
  to omit pruned connections (\ref{fig:plotprune1}), whereas they can
  be viewed as dashed lines by including the \code{prune\_col}
  argument (\ref{fig:plotprune2}).}
\label{fig:plotprune}
\end{figure}
